---
title: "R Notebook"
output: html_notebook
---

```{r}
library(lpSolve, quietly = TRUE)
library(kableExtra, quietly = TRUE)
```


1.

```{r}
obj_fun <- c(300,0,700,0,400,500,600,300,200,200,500,0,0,0,400,500,300,0)
#x11,x12,x13,x21,x22,x23,x31,x32,x33,x41,x42,x43,x51,x52,x53,x61,x62,x63
```

```{r}
const_LHS <- matrix(c(1,0,0,1,0,0,1,0,0,1,0,0,1,0,0,1,0,0, #s1 cap
                      0,1,0,0,1,0,0,1,0,0,1,0,0,1,0,0,1,0, #s2 cap
                      0,0,1,0,0,1,0,0,1,0,0,1,0,0,1,0,0,1, #s3 cap
                      1,1,1,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0, #a1
                      0,0,0,1,1,1,0,0,0,0,0,0,0,0,0,0,0,0, #a2
                      0,0,0,0,0,0,1,1,1,0,0,0,0,0,0,0,0,0, #a3
                      0,0,0,0,0,0,0,0,0,1,1,1,0,0,0,0,0,0, #a4
                      0,0,0,0,0,0,0,0,0,0,0,0,1,1,1,0,0,0, #a5
                      0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,1,1,1, #a6
                      0.02,0,0,0.07,0,0,0,0,0,-0.02,0,0,0.09,0,0,0.04,0,0, #s1 6th 30
                      -0.04,0,0,0.01,0,0,-0.06,0,0,-0.08,0,0,0.03,0,0,-0.02,0,0, #s1 6th 36
                      0.08,0,0,-0.02,0,0,0.02,0,0,0.1,0,0,0.04,0,0,-0.02,0,0, #s1 7th 30
                      0.02,0,0,-0.08,0,0,-0.04,0,0,0.04,0,0,-0.02,0,0,-0.08,0,0, #s1 7th 36
                      0,0,0,0.05,0,0,0.08,0,0,0.02,0,0,-0.03,0,0,0.08,0,0, #s1 8th 30
                      -0.06,0,0,-0.01,0,0,0.02,0,0,-0.04,0,0,-0.09,0,0,0.02,0,0, #s1 8th 36
                      0,0.02,0,0,0.07,0,0,0,0,0,-0.02,0,0,0.09,0,0,0.04,0, #s2 6th 30
                      0,-0.04,0,0,0.01,0,0,-0.06,0,0,-0.08,0,0,0.03,0,0,-0.02,0, #s2 6th 36
                      0,0.08,0,0,-0.02,0,0,0.02,0,0,0.1,0,0,0.04,0,0,-0.02,0, #s2 7th 30
                      0,0.02,0,0,-0.08,0,0,-0.04,0,0,0.04,0,0,-0.02,0,0,-0.08,0, #s2 7th 36
                      0,0,0,0,0.05,0,0,0.08,0,0,0.02,0,0,-0.03,0,0,0.08,0, #s2 8th 30
                      0,-0.06,0,0,-0.01,0,0,0.02,0,0,-0.04,0,0,-0.09,0,0,0.02,0, #s2 8th 36
                      0,0,0.02,0,0,0.07,0,0,0,0,0,-0.02,0,0,0.09,0,0,0.04, #s3 6th 30
                      0,0,-0.04,0,0,0.01,0,0,-0.06,0,0,-0.08,0,0,0.03,0,0,-0.02, #s3 6th 36
                      0,0,0.08,0,0,-0.02,0,0,0.02,0,0,0.1,0,0,0.04,0,0,-0.02, #s3 7th 30
                      0,0,0.02,0,0,-0.08,0,0,-0.04,0,0,0.04,0,0,-0.02,0,0,-0.08, #s3 7th 36
                      0,0,0,0,0,0.05,0,0,0.08,0,0,0.02,0,0,-0.03,0,0,0.08, #s3 8th 30
                      0,0,-0.06,0,0,-0.01,0,0,0.02,0,0,-0.04,0,0,-0.09,0,0,0.02, #s3 8th 36
                      0,0,0,1,0,0,0,0,0,0,0,0,0,0,0,0,0,0, #infeasible x21
                      0,0,0,0,0,0,0,0,0,0,0,1,0,0,0,0,0,0, #infeasible x43
                      0,0,0,0,0,0,0,0,0,0,0,0,0,1,0,0,0,0), #infeasible x52
                    byrow=TRUE, ncol=18)
```


```{r}
const_dir <- c("<=",
               "<=",
               "<=",
               "=",
               "=",
               "=",
               "=",
               "=",
               "=",
               ">=",
               "<=",
               ">=",
               "<=",
               ">=",
               "<=",
               ">=",
               "<=",
               ">=",
               "<=",
               ">=",
               "<=",
               ">=",
               "<=",
               ">=",
               "<=",
               ">=",
               "<=",
               "=",
               "=",
               "=")
```

```{r}
const_RHS <- c(900,
               1100,
               1000,
               450,
               600,
               550,
               350,
               500,
               450,
               0,
               0,
               0,
               0,
               0,
               0,
               0,
               0,
               0,
               0,
               0,
               0,
               0,
               0,
               0,
               0,
               0,
               0,
               0,
               0,
               0)
```


```{r}
LP_model <- lp(direction='min',
               objective.in = obj_fun,
               const.mat= const_LHS,
               const.dir = const_dir,
               const.rhs = const_RHS,
               compute.sens = TRUE,
               all.int = TRUE
) 
```

```{r}
LP_model
```

optimal solution has total cost of $555,600


```{r}
optim_val <- LP_model$solution
optim_val
```

optimal solution:
send all 450 from area 1 to school 2
send 432 kids from area 2 to school 2 and 168 to school 3
send 4 kids from area 3 to school 1, 218 to school 2, and 328 to school 3
send all 350 kids from area 4 to school 1
send 364 kids from area 5 to school 1 and 136 to school 3
send 82 kids from area 6 to school 1 and 368 to school 3

school 1 students = 0+0+4+350+364+82 = 800 - non binding capacity (slack = 100)
school 2 = 450+432+218+0+0+0 = 1100 - binding capacity
school 3 = 0+168+328+0+136+368 = 1000 - binding capacity


2a.

```{r}
sens_LB <- LP_model$sens.coef.from
sens_LB
```

```{r}
sens_UB <- LP_model$sens.coef.to
sens_UB
```

upper bound for area 6 school one is 680, increase is to 550, don't need to re-solve

2b.
 
 area 6 school 2: upper limit is infinite - don't need to re-solve won't affect solution, will still be 0 students
 area 6 school 3: busing cost is 0 so 10% increase would still be 0
 
 
 3a. (still confused how to interpret)
 
```{r}
LP_model$duals
```

```{r}
LP_model$duals.from
```
 
```{r}
LP_model$duals.to
```
 
 
 
 --
 EVERYTHING BELOW THIS IS FROM BEFORE WHEN WE HADN'T ADDED INT CONSTRAINT
 
 

```{r}
obj_fun2 <- c(300,0,700,0,400,500,600,300,200,200,500,0,0,0,400,550,300,0)
LP_model2 <- lp(direction='min',
               objective.in = obj_fun2,
               const.mat= const_LHS,
               const.dir = const_dir,
               const.rhs = const_RHS,
               compute.sens = TRUE
)
LP_model2
```

```{r}
optim_val2 <- LP_model2$solution
optim_val2
```

```{r}
sens_UB2 <- LP_model2$sens.coef.to
sens_UB2
```

no limit for area 6 school 2, don't need to re-solve 

find new optimal cost

```{r}
obj_fun3 <- c(300,0,700,0,400,500,600,300,200,200,500,0,0,0,400,550,330,0)
LP_model3 <- lp(direction='min',
               objective.in = obj_fun3,
               const.mat= const_LHS,
               const.dir = const_dir,
               const.rhs = const_RHS,
               compute.sens = TRUE
)
LP_model3
```

figure out slack



back to LP_model without bus increases for area 6
```{r}
shadow_prices <- LP_model$duals
shadow_prices
```

```{r}
LP_model$duals.to
```

```{r}
LP_model$duals.from
```

I have different values for this than them
are we interpreting the negative shadow prices as reducing costs or increasing costs (not sure if it's diff bc it's minimization)

if we increase students by one unit we will decrease (-) or increase (+) costs by x amount
increasing by more = multiply by the shadow price
look at if the cost will increase so much that it wouldn't be worth it or if it would increase so if you add a classroom you wouldn't decrease the cost
can't incrase infinitely - can increase up to the slack level, once you go below that you will need to re-solve the model 
so have to consider which is non binding 
slack should be 560 - can only decrease by that
make sure to use integer constraints with all.int = TRUE
consider each area separately

part 4

has to be a binary integer problem
decision variables need to be binary


```{r}
obj_fun4 <- c(300,0,700,0,400,500,600,300,200,200,500,0,0,0,400,500,300,0)
#x11,x12,x13,x21,x22,x23,x31,x32,x33,x41,x42,x43,x51,x52,x53,x61,x62,x63
```

```{r}
const_LHS4 <- matrix(c(1,0,0,1,0,0,1,0,0,1,0,0,1,0,0,1,0,0, #s1 cap
                      0,1,0,0,1,0,0,1,0,0,1,0,0,1,0,0,1,0, #s2 cap
                      0,0,1,0,0,1,0,0,1,0,0,1,0,0,1,0,0,1, #s3 cap
                      1,1,1,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0, #a1
                      0,0,0,1,1,1,0,0,0,0,0,0,0,0,0,0,0,0, #a2
                      0,0,0,0,0,0,1,1,1,0,0,0,0,0,0,0,0,0, #a3
                      0,0,0,0,0,0,0,0,0,1,1,1,0,0,0,0,0,0, #a4
                      0,0,0,0,0,0,0,0,0,0,0,0,1,1,1,0,0,0, #a5
                      0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,1,1,1, #a6
                      0.02,0,0,0.07,0,0,0,0,0,-0.02,0,0,0.09,0,0,0.04,0,0, #s1 6th 30
                      -0.04,0,0,0.01,0,0,-0.06,0,0,-0.08,0,0,0.03,0,0,-0.02,0,0, #s1 6th 36
                      0.08,0,0,-0.02,0,0,0.02,0,0,0.1,0,0,0.04,0,0,-0.02,0,0, #s1 7th 30
                      0.02,0,0,-0.08,0,0,-0.04,0,0,0.04,0,0,-0.02,0,0,-0.08,0,0, #s1 7th 36
                      0,0,0,0.05,0,0,0.08,0,0,0.02,0,0,-0.03,0,0,0.08,0,0, #s1 8th 30
                      -0.06,0,0,-0.01,0,0,0.02,0,0,-0.04,0,0,-0.09,0,0,0.02,0,0, #s1 8th 36
                      0,0.02,0,0,0.07,0,0,0,0,0,-0.02,0,0,0.09,0,0,0.04,0, #s2 6th 30
                      0,-0.04,0,0,0.01,0,0,-0.06,0,0,-0.08,0,0,0.03,0,0,-0.02,0, #s2 6th 36
                      0,0.08,0,0,-0.02,0,0,0.02,0,0,0.1,0,0,0.04,0,0,-0.02,0, #s2 7th 30
                      0,0.02,0,0,-0.08,0,0,-0.04,0,0,0.04,0,0,-0.02,0,0,-0.08,0, #s2 7th 36
                      0,0,0,0,0.05,0,0,0.08,0,0,0.02,0,0,-0.03,0,0,0.08,0, #s2 8th 30
                      0,-0.06,0,0,-0.01,0,0,0.02,0,0,-0.04,0,0,-0.09,0,0,0.02,0, #s2 8th 36
                      0,0,0.02,0,0,0.07,0,0,0,0,0,-0.02,0,0,0.09,0,0,0.04, #s3 6th 30
                      0,0,-0.04,0,0,0.01,0,0,-0.06,0,0,-0.08,0,0,0.03,0,0,-0.02, #s3 6th 36
                      0,0,0.08,0,0,-0.02,0,0,0.02,0,0,0.1,0,0,0.04,0,0,-0.02, #s3 7th 30
                      0,0,0.02,0,0,-0.08,0,0,-0.04,0,0,0.04,0,0,-0.02,0,0,-0.08, #s3 7th 36
                      0,0,0,0,0,0.05,0,0,0.08,0,0,0.02,0,0,-0.03,0,0,0.08, #s3 8th 30
                      0,0,-0.06,0,0,-0.01,0,0,0.02,0,0,-0.04,0,0,-0.09,0,0,0.02, #s3 8th 36
                      0,0,0,1,0,0,0,0,0,0,0,0,0,0,0,0,0,0, #infeasible x21
                      0,0,0,0,0,0,0,0,0,0,0,1,0,0,0,0,0,0, #infeasible x43
                      0,0,0,0,0,0,0,0,0,0,0,0,0,1,0,0,0,0), #infeasible x52
                    byrow=TRUE, ncol=18)
```


```{r}
const_dir4 <- c("<=",
               "<=",
               "<=",
               "=",
               "=",
               "=",
               "=",
               "=",
               "=",
               ">=",
               "<=",
               ">=",
               "<=",
               ">=",
               "<=",
               ">=",
               "<=",
               ">=",
               "<=",
               ">=",
               "<=",
               ">=",
               "<=",
               ">=",
               "<=",
               ">=",
               "<=",
               "=",
               "=",
               "=")
```

```{r}
const_RHS4 <- c(900,
               1100,
               1000,
               450,
               600,
               550,
               350,
               500,
               450,
               0,
               0,
               0,
               0,
               0,
               0,
               0,
               0,
               0,
               0,
               0,
               0,
               0,
               0,
               0,
               0,
               0,
               0,
               0,
               0,
               0)
```


```{r}
LP_model4 <- lp(direction='min',
               objective.in = obj_fun4,
               const.mat= const_LHS4,
               const.dir = const_dir4,
               const.rhs = const_RHS4,
               compute.sens = TRUE,
)
```

```{r}
LP_model4
```


```{r}
optim_val <- LP_model4$solution
optim_val
```
